results

### 2022-04-10 code/analysis_code/particle_type.Rmd
color updates, editing figures

```{r}
part_sum %>% ggplot(aes(x = name, y = value, fill = name)) + 
  geom_col() + geom_text(aes(label = percent(value/559), vjust = -0.5, position = position_dodge(width=0.9))) + ylim(0, 350) 
```

@@

>>>>

```{r}
part_sum %>% ggplot(aes(x = name, y = value, fill = name)) + 
  geom_col() + geom_text(aes(label = percent(value/559), vjust = -0.5)) + ylim(0,350) + guides(fill=guide_legend("Particle Type")) + labs(y = "Microplastic Composition")
#position = position_dodge(width=0.9))) + ylim(0, 350) 
```

@@

#particle color breakdown (aggregate, for July only)
```{r}
# load particle color data
color <- read_excel(here("data", "raw_data", "mp_color.xlsx"))

# plot color data
color %>% ggplot(aes(x = color, y = number, fill = color)) + 
  geom_col() + geom_text(aes(label = percent(number/968), vjust = -0.5)) + ylim(0,420) + scale_fill_manual(values = c("black" = "gray0", "blue" = "royalblue2", "green" = "seagreen2", "other" = "yellow1", "red" = "red1", "transparent/white"= "white")) + labs(y = "Occurence") + guides(fill=guide_legend("Particle Color"))

#stacked

#reorder
color$color <- reorder(color$color, color$number) #largest to smallest
# plot reordered stacked plot
ggplot(data = color, aes(x = obs, y = number, fill = color)) + geom_bar(stat = "identity", position = "fill") + scale_fill_manual(values = c("black" = "gray0", "blue" = "royalblue2", "green" = "seagreen2", "other" = "yellow1", "red" = "red1", "transparent/white"= "white")) + labs(x = "", y = "Microplastic Composition") + guides(fill=guide_legend("Particle Color")) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

 # geom_text(aes(label = paste0(percent(number/968))), position=position_stack(vjust=0.5), colour="white", size = 2)

 # here is some code for adding % labels, but I can't get it to work without messing up the graph                                                       #geom_text(aes(label=paste0(sprintf("%1.1f", percent*100),"%")),
                   #  position=position_stack(vjust=0.5), colour="white", size = 2)  
```

### 2022-04-10 code/analysis_code/results.Rmd
color updates, editing figures

#load packages and data
```{r}
library(ggplot2) #for plotting
library(broom) #for cleaning up output from lm()
library(here) #for data loading/saving
library(tidymodels) #for modeling
library(gridExtra) #for exporting tables
library(gridExtra) #for savings images in grids
library(dplyr)

@@

```{r}
#visualize linear model
lm_dup <- ggplot(all_counts_cv, aes(x = a_count, y = b_count)) + geom_point() + geom_smooth(method = lm) + labs(title = "MP Concentration Duplicate Counts", x = "A Count", y = "B Count")
lm_dup
```

@@

@@

```

>>>>

#load packages and data
```{r}
library(ggplot2) #for plotting
library(broom) #for cleaning up output from lm()
library(here) #for data loading/saving
library(tidymodels) #for modeling
library(gridExtra) #for exporting tables
library(gridExtra) #for savings images in grids
library(dplyr)
library(gridExtra) #for savings images in grids
library(gridExtra) #for savings images in grids

@@

```{r}
#visualize linear model
lm_dup <- ggplot(all_counts_cv, aes(x = a_count, y = b_count)) + geom_point() + geom_smooth(method = lm) + labs(title = "MP Concentration Duplicate Counts", x = "A Count", y = "B Count") + geom_abline(intercept = 0, slope = 1, colour='#E41A1C') + scale_x_log10() + scale_y_log10()
lm_dup
```

@@

We could use this same type of graph to display the comparative CV for my counts vs volunteer counts by doing AvsB for sample dates 1-3 and then AvsB for sample date 4 or AvsC and BvsD

#visualizing CV
```{r}
#visualize linear model
#subset data
all_counts_me <- all_counts_cv %>% filter(!date == "july_21") 
#plot
lm_me <- ggplot(all_counts_me, aes(x = a_count, y = b_count)) + geom_point() + geom_smooth(method = lm) + labs(title = "Duplicate Counts Nov, Apr, July", x = "A Count", y = "B Count") + xlim(0,25) + ylim(0,25) + geom_abline(intercept = 0, slope = 1, colour='#E41A1C')
lm_me #show plot

#subset data
all_counts_vol <- all_counts_cv %>% filter(date == "july_21")
#plot
lm_vol <- ggplot(all_counts_vol, aes(x = a_count, y = b_count)) + geom_point() + geom_smooth(method = lm) + labs(title = "Volunteer Counts July", x = "A Count", y = "B Count") + xlim(0,25) + ylim(0,25) + geom_abline(intercept = 0, slope = 1, colour='#E41A1C')
lm_vol #show plot
```

# CV for each sample date
```{r}
#nov
nov <- all_counts_cv %>% filter(date == "nov_20")
#plot
lm_nov <- ggplot(nov, aes(x = a_count, y = b_count)) + geom_point() + geom_smooth(method = lm) + labs(title = "Nov", x = "A Count", y = "B Count") + scale_x_log10()+ scale_y_log10() + geom_abline(intercept = 0, slope = 1, colour='#E41A1C') + expand_limits(x = 0, y = 0) 
lm_nov #show plot

#feb
feb <- all_counts_cv %>% filter(date == "feb_21")
#plot
lm_feb <- ggplot(feb, aes(x = a_count, y = b_count)) + geom_point() + geom_smooth(method = lm) + labs(title = "Feb", x = "A Count", y = "B Count") + scale_x_log10()+ scale_y_log10() + geom_abline(intercept = 0, slope = 1, colour='#E41A1C') 
lm_feb #show plot

#april
april <- all_counts_cv %>% filter(date == "apr_21")
#plot
lm_apr <- ggplot(april, aes(x = a_count, y = b_count)) + geom_point() + geom_smooth(method = lm) + labs(title = "April", x = "A Count", y = "B Count") + scale_x_log10()+ scale_y_log10() + geom_abline(intercept = 0, slope = 1, colour='#E41A1C') 
lm_apr #show plot

#july
all_counts_vol <- all_counts_cv %>% filter(date == "july_21")
#plot
lm_vol <- ggplot(all_counts_vol, aes(x = a_count, y = b_count)) + geom_point() + geom_smooth(method = lm) + labs(title = "July", x = "A Count", y = "B Count") + ylim(0,25) + scale_x_log10()+ scale_y_log10() + geom_abline(intercept = 0, slope = 1, colour='#E41A1C') + xlim(3,25) + ylim(3,30)
lm_vol #show plot

bivar_grid <- grid.arrange(lm_nov, lm_feb, lm_apr, lm_vol, ncol = 2)
bivar_grid #show grid
```

@@

```
# main rivers vs tributaries
```{r}
river <- sampledata %>% filter(watershed == "North Oconee River"|watershed == "Middle Oconee River"|watershed == "Oconee River")

trib <- sampledata %>% filter(!watershed == "North Oconee River", !watershed == "Middle Oconee River", !watershed == "Oconee River")

river <- river %>% mutate(body = "river")
trib <- trib %>% mutate(body = "trib")

#rejoin
body <- full_join(river, trib)

#plot
body %>% ggplot(aes(x = body, y = particles_l, color = )) + 
  geom_boxplot()  +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  labs(title = "Microplastic Concentration by Water Body", x = "Water Body", y = "Particles/Liter") + ylim(0,500) + scale_y_log10()

#kruskal wallis/mann whitney test - nonparametric test
kruskal.test(particles_l ~ body, data = body)
```

P=value less than 0.05 - we conclude that there IS a significant difference between groups. Let's do a pairwise.wilcox.test() to double check. 
# pairwise wilcox test to see differences between groups
```{r}
pairwise.wilcox.test(body$particles_l, body$body,
                 p.adjust.method = "BH")
```

This tells us that there is a significant difference between the MP concentrations in rivers and tributaries! Cool! We can try to tweak this boxplot figure to display those statistics, perhaps.
